### 2019-03-22 paper/sex_selection_spacing-ver2.tex
Updata regional results in Conclusion; explain periods used in Introduction

Finally, how are the changes in birth intervals with increasing access to sex selection 
distributed geographically?
Given the size and cultural differences within India, it is unlikely that the effect of
increasing access to sex selection will have had the same effect throughout 
\citep{retherford03b}.
I, therefore, examine how birth intervals have changed across four broad regions.
[TK only some groups]

TK How I am answering them

Direct information on the use of sex selection is not available, but it is possible
to compare different periods as a proxy since there has been substantial changes in access 
and legality of prenatal sex determination in India.
Abortion has been legal in India since 1971 and still is.
The first reports of sex determination appeared around 
1982--83 \citep{Sudha1999,bhat06,Grover2006}.
The number of clinics quickly increased, and knowledge about sex selection 
became widespread after a senior government official's wife aborted a 
fetus that turned out to be male \citep[p.\ 598]{Sudha1999}.
In 1994, the Central Government passed the Prenatal Diagnostic Techniques 
(PNDT) Act, making determining and communicating the sex of a fetus illegal.%
\footnote{
Details about the act are at \href{http://pndt.gov.in/}{http://pndt.gov.in/}.
The number of convictions has been low.
It took until January 2008 for the first state, Haryana, to reach five convictions.
Hence, private clinics apparently operate with little risk of legal action 
\citep{Sudha1999}.
Furthermore, there is little evidence that bans like this significantly
affect sex ratios \citep{Das-Gupta2016}.
}
[TK something on split 1995 to 2016 into two groups?]

@@

Three major, interrelated trends emerge from the data over the four decades.
First, fertility has declined for all groups, as shown by the parity progression
probabilities, with a larger decline the more educated the woman.
Across all groups, the likelihood of having an additional child in each spell depends 
strongly on how many sons you already have, with women with none or only one son having
the highest probability of a birth.
Second, there has been a general increase in the length of average birth 
intervals for all education groups.
The increase in average birth interval is larger the higher the spell and the higher the 
education group.
Finally, as shown by the increasing sex ratios for women with no or only one son there
has been a substantial increase in the use of sex selection over time.
As for fertility and birth intervals, this increase is larger the higher the spell and
with higher education.

@@

TK Regional differences

>>>>

Finally, how are the changes in birth intervals with increasing access to sex selection 
distributed geographically?
Given the size and cultural differences within India, it is unlikely that the effect of
increasing access to sex selection will have had the same effect throughout 
\citep{retherford03b}.
I, therefore, examine how birth intervals have changed across four broad regions.

Direct information on the use of sex selection is not available, but it is possible
to compare different periods as a proxy since there has been substantial changes in access 
and legality of prenatal sex determination in India.
Abortion has been legal in India since 1971 and still is.
The first reports of sex determination appeared around 1982--83 
\citep{Sudha1999,bhat06,Grover2006}.
The number of clinics quickly increased, and knowledge about sex selection became 
widespread after a senior government official's wife aborted a fetus that turned out to be 
male \citep[p.\ 598]{Sudha1999}.
In 1994, the Central Government passed the Prenatal Diagnostic Techniques (PNDT) Act, 
making determining and communicating the sex of a fetus illegal.%
\footnote{
Details about the act are at \href{http://pndt.gov.in/}{http://pndt.gov.in/}.
The number of convictions has been low.
It took until January 2008 for the first state, Haryana, to reach five convictions.
Hence, private clinics apparently operate with little risk of legal action 
\citep{Sudha1999}.
Furthermore, there is little evidence that bans like this significantly
affect sex ratios \citep{Das-Gupta2016}.
}

Based on these development, I split the four decades into four periods.
The first period cover until prenatal sex determination became widely available, 
1972--1984.
The second period saw (mostly) unrestricted access to sex selection, 1985--1994.
Although, as reviewed above, there is evidence that there has been a substantial 
increase in the use of sex selection, some researchers argue that we might be close to or 
have already passed a turning-point in the use of sex selection 
\citep{Das_Gupta2009,Diamond-Smith2015}.
I, therefore, split the period from the implementation of the PNDT Act until the last 
survey into two, 1995--2004, and 2005--2016.

@@

Three major, interrelated trends over the the four decades emerge from the data.
First, fertility has declined for all groups, as shown by the parity progression
probabilities, with a larger decline the more educated the woman.
Across all groups, the likelihood of having an additional child in each spell depends 
strongly on how many sons you already have, with women with none or only one son having
the highest probability of a birth.
Second, there has been a general increase in the length of average birth 
intervals for all education groups.
The increase in average birth interval is larger the higher the spell and the higher the 
education group.
Finally, as shown by the increasing sex ratios for women with no or only one son there
has been a substantial increase in the use of sex selection over time.
As for fertility and birth intervals, this increase is larger the higher the spell and
with higher education.

@@

With a country as large as India, it is natural to expect variation in both fertility
behavior, spacing patterns, and the degree of son preference.
The regional results confirm this, although differences are less apparent than one would
expect given the prior research on regional differences.
The ``North'' and the ``West'' regions have relatively similar spacing patterns, likely
precipitated by rapidly declining fertility and substantial use of sex selection.
The exception is that the rural areas of the ``North'' still has higher fertility and
consequently less use of sex selection and have, therefore, not yet experienced the
same reversal of spacing patterns.
Even the ``South'', which is otherwise considered the area with the least son preference,
has has seen a substantial increase in the use of sex selection and the associated changes 
in the spacing patterns.
Even though the ``South'' is often mentioned as the low-fertility region of India, 
fertility has declined fastest in the ``East'' region, although this does not appear to 
have translated into the use of sex selection.
